---
title: "Herbicide - 03: Alpha Diversity"
author: "Jan Waelchli"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

\pagebreak

```{r setup, include=FALSE, echo=F, warning=F, message=F,}

##clear the object from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

## set source to file location
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))

## load functions
source("functions/functions.R")
functions(path="functions/")

## install (if necessary) and load libraries
libraries()

```

```{r import, echo=F, message=F, warning=F}

# Import

# set wd to RDS files
setwd("interim/01_Import_Normalization")

#cImport RDS files
bDESIGN <- readRDS("bDESIGN.RDS")
bDAT <- readRDS("bDAT.RDS")
bDAT_rare <- readRDS("bDAT_rare.RDS")
b_rare <- readRDS("b_rare.RDS")

fDESIGN <- readRDS("fDESIGN.RDS")
fDAT <- readRDS("fDAT.RDS")
fDAT_rare <- readRDS("fDAT_rare.RDS")
f_rare <- readRDS("f_rare.RDS")

level_cols_Herbicide <- readRDS("level_cols_Herbicide.RDS")
level_cols_Application <- readRDS("level_cols_Application.RDS")

```
   

# Alpha diversity

We answered the following questions for the bacterial and fungal alpha diversity:  

* **Q1: Are there differences in alpha diversity between the herbicide applications?**
* **Q2: Are there differences in alpha diversity between the herbicide treatments?**

\vspace{10mm}

## Rarefaction

Before analyzing the alpha diversity, we conducted the rarefaction plot to make sure that no diversity was lost due to a too low rarefaction threshold.

### Figure 3 | Rarefaction

\vspace{5mm}

```{r rarefaction, eval=T, echo=F, message=F, warning=F, fig.height=4, fig.width=10}

#BACTERIA

#rarefy
pdf("/dev/null") #prevents drawing of standard rarefaction plot (we will create our own with colors)
brarefaction <- vegan::rarecurve(t(bDAT), step=200, 7000, label = FALSE)
names(brarefaction) <- bDESIGN$Label
invisible(dev.off())

# long-transformation.
protox <- mapply(FUN = function(x, y) {
  mydf <- as.data.frame(x)
  colnames(mydf) <- "value"
  mydf$species <- y
  mydf$subsample <- attr(x, "Subsample")
  mydf
}, x = brarefaction, y = as.list(names(brarefaction)), SIMPLIFY = FALSE)

brarefaction_long <- do.call(rbind, protox)
rownames(brarefaction_long) <- NULL  # pretty

#add Experiment
brarefaction_long$Experiment <- NA
for (i in 1:nrow(brarefaction_long)) {
  brarefaction_long$Experiment[i] <- as.character(bDESIGN$Experiment[bDESIGN$Label == brarefaction_long$species[i]])
}

#add Herbicide
brarefaction_long$Herbicide <- NA
for (i in 1:nrow(brarefaction_long)) {
  brarefaction_long$Herbicide[i] <- as.character(bDESIGN$Herbicide[bDESIGN$Label == brarefaction_long$species[i]])
}

#plot
bac_rareplot <- ggplot(brarefaction_long, aes(x = subsample, y = value, group = species)) +
                  theme_bw() +
                  geom_line(aes(color=Herbicide, lty=Experiment))+
                  scale_color_manual(values = level_cols_Herbicide)+
                  xlab("Sequencing Depth")+
                  ylab("Number of ASVs")+
                  ggtitle("Bacteria: Rarefaction")+
                  geom_vline(xintercept = b_rare, lty=2, lwd=1, col="dimgrey")

#FUNGI

#rarefy
pdf("/dev/null") #prevents drawing of standard rarefaction plot (we will create our own with colors)
frarefaction <- vegan::rarecurve(t(fDAT), step=20, 7000, label = FALSE)
names(frarefaction) <- fDESIGN$Label
invisible(dev.off())

# long-transformation.
protox <- mapply(FUN = function(x, y) {
  mydf <- as.data.frame(x)
  colnames(mydf) <- "value"
  mydf$species <- y
  mydf$subsample <- attr(x, "Subsample")
  mydf
}, x = frarefaction, y = as.list(names(frarefaction)), SIMPLIFY = FALSE)

frarefaction_long <- do.call(rbind, protox)
rownames(frarefaction_long) <- NULL  # pretty

#add Experiment
frarefaction_long$Experiment <- NA
for (i in 1:nrow(frarefaction_long)) {
  frarefaction_long$Experiment[i] <- as.character(fDESIGN$Experiment[fDESIGN$Label == frarefaction_long$species[i]])
}

#add Herbicide
frarefaction_long$Herbicide <- NA
for (i in 1:nrow(frarefaction_long)) {
  frarefaction_long$Herbicide[i] <- as.character(fDESIGN$Herbicide[fDESIGN$Label == frarefaction_long$species[i]])
}

#plot
fun_rareplot <- ggplot(frarefaction_long, aes(x = subsample, y = value, group = species)) +
                  theme_bw() +
                  geom_line(aes(color=Herbicide, lty=Experiment))+
                  scale_color_manual(values = level_cols_Herbicide)+
                  xlab("Sequencing Depth")+
                  ylab("Number of ASVs")+
                  ggtitle("Fungi: Rarefaction")+
                  geom_vline(xintercept = f_rare, lty=2, lwd=1, col="dimgrey")

#combine the two plots (library cowplot)
plot <- plot_grid(bac_rareplot + theme(legend.position="none"),
               fun_rareplot + theme(legend.position="none") + labs(colour=""),
               #labels=c("A","B"),
               align = "hv",
               nrow=1)

legend <- get_legend(bac_rareplot + theme(legend.position="bottom") + labs(fill="Experiment"))
fig4 <- plot_grid(plot, legend, nrow=2, rel_heights=c(1, .1))

#print
#png("figures/fig4.png",7000, 3200, res = 600)
fig4
#dev.off()

```

\vspace{5mm}

**Conclusion:** We chose rarefaction thresholds that do not affect the alpha diversity.

\pagebreak

## Method

We rarefied the dataset by the sequencing depth of `r b_rare` for bacteria and `r f_rare` for fungi and calculated the Shannon diversity as a index for alpha diversity for each sample. This was repeated 100 times. Then, the mean value from the 100 iterations was taken for statistical analysis between the different samples.  

```{r a-div shannon, warning=F, echo=F, message=F, eval=T}

### BACTERIA ###

repetitions <- 100

  #rarefy data a 100 time and calculate each time the alpha diversity (shannon)
  #differs each time because rarefaction is a random process
  diversity_bDAT_mat <- c()
  for (i in 1:repetitions){
    #generate rarefied community
    set.seed(i)#change seed every round, otherwise rrarefy will always have the same output
    temp_mat <- t(rrarefy(t(bDAT), b_rare))
    #calculate shannon
    shannon <- vegan::diversity(temp_mat, index="shannon", MARGIN=2)
    d <- exp(shannon)
    #combine
    diversity_bDAT_mat <- cbind(diversity_bDAT_mat, d)
  }

#calculate means of subsamples
asv_diversity <- rowMeans(diversity_bDAT_mat)
#put them all in one dataframe
alpha_diversity_bDAT_summary <- cbind(asv_diversity, bDESIGN)


### FUNGI ###

  #rarefy data a 100 time and calculate each time the alpha diversity (shannon)
  #differs each time because rarefaction is a random process
  diversity_fDAT_mat <- c()
  for (i in 1:repetitions){
    #generate rarefied community
    temp_mat <- t(rrarefy(t(fDAT), f_rare))
    #calculate shannon
    shannon <- vegan::diversity(temp_mat, index="shannon", MARGIN=2)
    d <- exp(shannon)
    #combine
    diversity_fDAT_mat <- cbind(diversity_fDAT_mat, d)
  }

#calculate means of subsamples
asv_diversity <- rowMeans(diversity_fDAT_mat)
#put them all in one dataframe
alpha_diversity_fDAT_summary <- cbind(asv_diversity, fDESIGN)


```

## Effect of all factors on alpha diversity

Before answering the questions, we got an overview for both experiments by investigating the effect on alpha diversity by the factors of different herbicides, different herbicide applications, time, any of their interactions and technical lab artefacts (the DNA extraction and PCR amplification was done on two plates). We modeled the alpha diversity against these factors in an aov-model and performed a F-Test.

### Bacteria

```{r a-div all bacteria tp * hc in exp1, warning=F, echo=F, message=F, eval=T}

### BACTERIA

#split dataset by experiments
exp1_alpha_diversity_bDAT_summary <- droplevels(alpha_diversity_bDAT_summary[alpha_diversity_bDAT_summary$Experiment==1,])
exp1_bDESIGN <- droplevels(bDESIGN[bDESIGN$Experiment==1,])

exp2_alpha_diversity_bDAT_summary <- droplevels(alpha_diversity_bDAT_summary[alpha_diversity_bDAT_summary$Experiment==2,])
exp2_bDESIGN <- droplevels(bDESIGN[bDESIGN$Experiment==2,])

#all
#F test for sample groups by Application
exp1_balpha_all_aov <- aov(asv_diversity ~ Herbicide * Application * Time + Plate, data=exp1_alpha_diversity_bDAT_summary)
pander(summary(exp1_balpha_all_aov), caption = "Bacteria, experiment 1: F test")

exp2_balpha_all_aov <- aov(asv_diversity ~ Herbicide * Application * Time + Plate, data=exp2_alpha_diversity_bDAT_summary)
pander(summary(exp2_balpha_all_aov), caption = "Bacteria, experiment 2: F test")

```

\pagebreak

### Fungi

```{r a-div all fungi tp * hc in exp1, warning=F, echo=F, message=F, eval=T}

### FUNGI

#split dataset by experiments
exp1_alpha_diversity_fDAT_summary <- droplevels(alpha_diversity_fDAT_summary[alpha_diversity_fDAT_summary$Experiment==1,])
exp1_fDESIGN <- droplevels(fDESIGN[fDESIGN$Experiment==1,])

exp2_alpha_diversity_fDAT_summary <- droplevels(alpha_diversity_fDAT_summary[alpha_diversity_fDAT_summary$Experiment==2,])
exp2_fDESIGN <- droplevels(fDESIGN[fDESIGN$Experiment==2,])

#all
#F test for sample groups by Application
exp1_falpha_all_aov <- aov(asv_diversity ~ Herbicide * Application * Time + Plate, data=exp1_alpha_diversity_fDAT_summary)
pander(summary(exp1_falpha_all_aov), caption = "Fungi, experiment 1: F test")

exp2_falpha_all_aov <- aov(asv_diversity ~ Herbicide * Application * Time + Plate, data=exp2_alpha_diversity_fDAT_summary)
pander(summary(exp2_falpha_all_aov), caption = "Fungi, experiment 2: F test")


#export for paper

#combine tables
df1 <- cbind(Taxa="Bacteria", Experiment="Experiment1", Factor=rownames(summary(exp1_balpha_all_aov)[[1]]), summary(exp1_balpha_all_aov)[[1]])
df2 <- cbind(Taxa="Bacteria", Experiment="Experiment2", Factor=rownames(summary(exp2_balpha_all_aov)[[1]]), summary(exp2_balpha_all_aov)[[1]])
df3 <- cbind(Taxa="Fungi", Experiment="Experiment1", Factor=rownames(summary(exp1_falpha_all_aov)[[1]]), summary(exp1_falpha_all_aov)[[1]])
df4 <- cbind(Taxa="Fungi", Experiment="Experiment2", Factor=rownames(summary(exp2_falpha_all_aov)[[1]]), summary(exp2_falpha_all_aov)[[1]])
df_all <- rbind(df1, df2, df3, df4)
rownames(df_all) <- NULL
rm(df1, df2, df3, df4)

#save
#write.csv(df_all, "paper_tables/alpha_diversity.csv", row.names = F)



```


**Conclusion:** We see a strong effect on the alpha diversity between the two plates for bacteria and fungi. Samples were randomized on the two plates. We are not interested in differences between the two plates, that's why we kept the "plate-treatment" in the following models as a factor so that effects of the other factors were quantified while accounting for the plate variance. In bacteria, the alpha diversity changed over time. Herbicide application on different targets and using different herbicides did not significantly influence the alpha diversity. For fungi, applying herbicides either on weeds or direct on soil affected the alpha diversity while the time nor the herbicide had a significant influence.

\pagebreak

## Application effect

**Q1: Are there differences in alpha diversity between the herbicides applications?**  
The herbicide was either applied directly on soil or on weeds. While we couldn't find any effect on the alpha diversity in bacterial communities, we found differences for fungi. We further investigated for both experiments how the different herbicides applications did shape the communities. For consistency we investigated the bacterial and fungal communities.

### Bacteria

```{r a-div Application effect bac, warning=F, echo=F, message=F, eval=T}

### BACTERIA ###

## Experiment 1 ##

### pair-wise tests
alpha_Application_aov_tukey <- emmeans::emmeans(exp1_balpha_all_aov, "Application")  # library(emmeans)
alpha_Application_aov_tukey_letters <- multcomp::cld(alpha_Application_aov_tukey, Letter="abcdefghi", alpha=0.05)
row.names(alpha_Application_aov_tukey_letters) <- alpha_Application_aov_tukey_letters[,1]
alpha_Application_aov_tukey_letters <- alpha_Application_aov_tukey_letters[levels(exp1_bDESIGN$Application),]
# means
alpha_Application_means <- aggregate(exp1_alpha_diversity_bDAT_summary$asv_diversity,list(exp1_alpha_diversity_bDAT_summary$Application), mean)
row.names(alpha_Application_means) <- alpha_Application_means[,1]
alpha_Application_means <- alpha_Application_means[levels(exp1_bDESIGN$Application),]
# table
alpha_table_Application <- cbind(Shannon=alpha_Application_means[,2], alpha_Application_aov_tukey_letters[,2:7])
alpha_table_Application <- alpha_table_Application[,c(1,7)]

pander(alpha_table_Application, caption = "Bacteria: Application effect in experiment 1")


## Experiment 2 ##

### pair-wise tests
alpha_Application_aov_tukey <- emmeans::emmeans(exp2_balpha_all_aov, "Application")  # library(emmeans)
alpha_Application_aov_tukey_letters <- multcomp::cld(alpha_Application_aov_tukey, Letter="abcdefghi", alpha=0.05)
row.names(alpha_Application_aov_tukey_letters) <- alpha_Application_aov_tukey_letters[,1]
alpha_Application_aov_tukey_letters <- alpha_Application_aov_tukey_letters[levels(exp2_bDESIGN$Application),]
# means
alpha_Application_means <- aggregate(exp2_alpha_diversity_bDAT_summary$asv_diversity,list(exp2_alpha_diversity_bDAT_summary$Application), mean)
row.names(alpha_Application_means) <- alpha_Application_means[,1]
alpha_Application_means <- alpha_Application_means[levels(exp2_bDESIGN$Application),]
# table
alpha_table_Application <- cbind(Shannon=alpha_Application_means[,2], alpha_Application_aov_tukey_letters[,2:7])
alpha_table_Application <- alpha_table_Application[,c(1,7)]

pander(alpha_table_Application, caption = "Bacteria: Application effect in experiment 2")

```

### Fungi

```{r a-div Application effect fungi, warning=F, echo=F, message=F, eval=T}

### FUNGI ###

## Experiment 1 ##

### pair-wise tests
alpha_Application_aov_tukey <- emmeans::emmeans(exp1_falpha_all_aov, "Application")  # library(emmeans)
alpha_Application_aov_tukey_letters <- multcomp::cld(alpha_Application_aov_tukey, Letter="abcdefghi", alpha=0.05)
row.names(alpha_Application_aov_tukey_letters) <- alpha_Application_aov_tukey_letters[,1]
alpha_Application_aov_tukey_letters <- alpha_Application_aov_tukey_letters[levels(exp1_fDESIGN$Application),]
# means
alpha_Application_means <- aggregate(exp1_alpha_diversity_fDAT_summary$asv_diversity,list(exp1_alpha_diversity_fDAT_summary$Application), mean)
row.names(alpha_Application_means) <- alpha_Application_means[,1]
alpha_Application_means <- alpha_Application_means[levels(exp1_fDESIGN$Application),]
# table
alpha_table_Application <- cbind(Shannon=alpha_Application_means[,2], alpha_Application_aov_tukey_letters[,2:7])
alpha_table_Application <- alpha_table_Application[,c(1,7)]

pander(alpha_table_Application, caption = "Fungi: Application effect in experiment 1")


## Experiment 2 ##

### pair-wise tests
alpha_Application_aov_tukey <- emmeans::emmeans(exp2_falpha_all_aov, "Application")  # library(emmeans)
alpha_Application_aov_tukey_letters <- multcomp::cld(alpha_Application_aov_tukey, Letter="abcdefghi", alpha=0.05)
row.names(alpha_Application_aov_tukey_letters) <- alpha_Application_aov_tukey_letters[,1]
alpha_Application_aov_tukey_letters <- alpha_Application_aov_tukey_letters[levels(exp2_fDESIGN$Application),]
# means
alpha_Application_means <- aggregate(exp2_alpha_diversity_fDAT_summary$asv_diversity,list(exp2_alpha_diversity_fDAT_summary$Application), mean)
row.names(alpha_Application_means) <- alpha_Application_means[,1]
alpha_Application_means <- alpha_Application_means[levels(exp2_fDESIGN$Application),]
# table
alpha_table_Application <- cbind(Shannon=alpha_Application_means[,2], alpha_Application_aov_tukey_letters[,2:7])
alpha_table_Application <- alpha_table_Application[,c(1,7)]

pander(alpha_table_Application, caption = "Fungi: Application effect in experiment 2")

```

### Figure 4.1 | Shannon diversity - Application

\vspace{5mm}

```{r plot a-div Application, echo=F, message=F, warning=F, eval=T, fig.height=5, fig.width=8}

### BACTERIA

#data frame for letters
letters <- data.frame(label=c(rep("a",4)),
                      Application=rep(c("weeds", "soil"), 2),
                      Experiment=rep(c("Experiment 1","Experiment 2"),each=2),
                      x=c(rep(c(1:2),2)),
                      y=c(800,800,800,800))

alpha_diversity_bDAT_summary$Time <- as.factor(alpha_diversity_bDAT_summary$Time)
levels(alpha_diversity_bDAT_summary$Experiment) <- c("Experiment 1", "Experiment 2")


#plot
bac_plot <- ggplot(alpha_diversity_bDAT_summary, aes(y=asv_diversity, x=Application, fill=Application)) +
              geom_boxplot(position=position_dodge2(width=0.75, preserve="single"),outlier.colour = NA, alpha=0.7) +
              geom_jitter(aes(shape=Time), size=2, position=position_jitterdodge(jitter.width=0.7,dodge.width = 0.75), alpha=0.3)+
              theme_bw() +
              theme(legend.position="none") +
              ylab("Shanon Diversity")+
              xlab("")+
              scale_fill_manual(values=level_cols_Application) +
              scale_shape_manual(values=c(15,17)) +
              facet_grid(.~ Experiment) +
              labs(title = "Bacteria: Alpha Diversity", subtitle = "Alpha Diversity ~ Herbicide * Application * Time + Plate")+
              #labs(subtitle = "Alpha Diversity ~ Herbicide * Application * Time + Plate")+
              theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.subtitle=element_text(size=10))+
              geom_text(data = letters, aes(x = x, y = y, label = label, Application=Application),size=5, col="dimgrey")


### FUNGI

#data frame for letters
letters <- data.frame(label=c("a", "b", "b", "a"),
                      Application=rep(c("weeds", "soil"), 2),
                      Experiment=rep(c("Experiment 1","Experiment 2"),each=2),
                      x=c(rep(c(1:2),2)),
                      y=c(58,58,58,58))

alpha_diversity_fDAT_summary$Time <- as.factor(alpha_diversity_fDAT_summary$Time)
levels(alpha_diversity_fDAT_summary$Experiment) <- c("Experiment 1", "Experiment 2")


#plot
fun_plot <- ggplot(alpha_diversity_fDAT_summary, aes(y=asv_diversity, x=Application, fill=Application)) +
              geom_boxplot(position=position_dodge2(width=0.75, preserve="single"),outlier.colour = NA, alpha=0.7) +
              geom_jitter(aes(shape=Time), size=2, position=position_jitterdodge(jitter.width=0.7,dodge.width = 0.75), alpha=0.3)+
              theme_bw() +
              theme(legend.position="none") +
              ylab("Shanon Diversity")+
              xlab("")+
              scale_fill_manual(values=level_cols_Application) +
              scale_shape_manual(values=c(15,17)) +
              facet_grid(.~ Experiment) +
              labs(title = "Fungi: Alpha Diversity", subtitle = "Alpha Diversity ~ Herbicide * Application * Time + Plate")+
              #labs(subtitle = "Alpha Diversity ~ Herbicide * Application * Time + Plate")+
              theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.subtitle=element_text(size=10))+
              geom_text(data = letters, aes(x = x, y = y, label = label, Application=Application),size=5, col="dimgrey")



### PLOT

#combine the two plots (library cowplot)
fig4.1 <- plot_grid(bac_plot + theme(legend.position="none"),
               fun_plot + theme(legend.position="none") + labs(colour=""),
               #labels=c("A","B"),
               align = "hv",
               nrow=1)

#for paper
#saveRDS(fig4.1, "paper_figs/RDS/fig4.1.RDS")

#add legend
legend <- get_legend(bac_plot + theme(legend.position="bottom") + labs(fill="Application"))
fig4.1 <- plot_grid(fig4.1, legend, nrow=2, rel_heights=c(1, .1))

#print
#png("figures/fig4.1.png",6400, 4000, res = 600)
fig4.1
#dev.off()


```

\vspace{5mm}

**Conclusion:** As already seen, bacterial alpha diversity remains constant between the two herbicide applications. In experiment 1 led applying herbicides on weeds to a higher fungal alpha diversity while we observed the opposite effect in experiment 2.

\vspace{5mm}

## Herbicide effect

**Q2: Are there differences in alpha diversity between the herbicide treatments?**  
We treated samples either with water (control), glyphosate or terbuthylazine. We saw above that the different herbicides did not influence the bacterial nor fungal alpha diversity.

```{r a-div Herbicide effect bac, warning=F, echo=F, message=F, eval=T}

# ### BACTERIA ###
# 
# ## Experiment 1 ##
# 
# ### pair-wise tests
# alpha_Herbicide_aov_tukey <- emmeans(exp1_balpha_all_aov, "Herbicide")  # library(emmeans)
# alpha_Herbicide_aov_tukey_letters <- cld(alpha_Herbicide_aov_tukey, Letter="abcdefghi", alpha=0.05)
# row.names(alpha_Herbicide_aov_tukey_letters) <- alpha_Herbicide_aov_tukey_letters[,1]
# alpha_Herbicide_aov_tukey_letters <- alpha_Herbicide_aov_tukey_letters[levels(exp1_bDESIGN$Herbicide),]
# # means
# alpha_Herbicide_means <- aggregate(exp1_alpha_diversity_bDAT_summary$asv_diversity,list(exp1_alpha_diversity_bDAT_summary$Herbicide), mean)
# row.names(alpha_Herbicide_means) <- alpha_Herbicide_means[,1]
# alpha_Herbicide_means <- alpha_Herbicide_means[levels(exp1_bDESIGN$Herbicide),]
# # table
# alpha_table_Herbicide <- cbind(Shannon=alpha_Herbicide_means[,2], alpha_Herbicide_aov_tukey_letters[,2:7])
# alpha_table_Herbicide <- alpha_table_Herbicide[,c(1,7)]
# 
# pander(alpha_table_Herbicide, caption = "Bacteria: Herbicide effect in experiment 1")
# 
# 
# ## Experiment 2 ##
# 
# ### pair-wise tests
# alpha_Herbicide_aov_tukey <- emmeans(exp2_balpha_all_aov, "Herbicide")  # library(emmeans)
# alpha_Herbicide_aov_tukey_letters <- cld(alpha_Herbicide_aov_tukey, Letter="abcdefghi", alpha=0.05)
# row.names(alpha_Herbicide_aov_tukey_letters) <- alpha_Herbicide_aov_tukey_letters[,1]
# alpha_Herbicide_aov_tukey_letters <- alpha_Herbicide_aov_tukey_letters[levels(exp2_bDESIGN$Herbicide),]
# # means
# alpha_Herbicide_means <- aggregate(exp2_alpha_diversity_bDAT_summary$asv_diversity,list(exp2_alpha_diversity_bDAT_summary$Herbicide), mean)
# row.names(alpha_Herbicide_means) <- alpha_Herbicide_means[,1]
# alpha_Herbicide_means <- alpha_Herbicide_means[levels(exp2_bDESIGN$Herbicide),]
# # table
# alpha_table_Herbicide <- cbind(Shannon=alpha_Herbicide_means[,2], alpha_Herbicide_aov_tukey_letters[,2:7])
# alpha_table_Herbicide <- alpha_table_Herbicide[,c(1,7)]
# 
# pander(alpha_table_Herbicide, caption = "Bacteria: Herbicide effect in experiment 2")

```

```{r a-div Herbicide effect fungi, warning=F, echo=F, message=F, eval=T}

# ### FUNGI ###
# 
# ## Experiment 1 ##
# 
# ### pair-wise tests
# alpha_Herbicide_aov_tukey <- emmeans(exp1_falpha_all_aov, "Herbicide")  # library(emmeans)
# alpha_Herbicide_aov_tukey_letters <- cld(alpha_Herbicide_aov_tukey, Letter="abcdefghi", alpha=0.05)
# row.names(alpha_Herbicide_aov_tukey_letters) <- alpha_Herbicide_aov_tukey_letters[,1]
# alpha_Herbicide_aov_tukey_letters <- alpha_Herbicide_aov_tukey_letters[levels(exp1_fDESIGN$Herbicide),]
# # means
# alpha_Herbicide_means <- aggregate(exp1_alpha_diversity_fDAT_summary$asv_diversity,list(exp1_alpha_diversity_fDAT_summary$Herbicide), mean)
# row.names(alpha_Herbicide_means) <- alpha_Herbicide_means[,1]
# alpha_Herbicide_means <- alpha_Herbicide_means[levels(exp1_fDESIGN$Herbicide),]
# # table
# alpha_table_Herbicide <- cbind(Shannon=alpha_Herbicide_means[,2], alpha_Herbicide_aov_tukey_letters[,2:7])
# alpha_table_Herbicide <- alpha_table_Herbicide[,c(1,7)]
# 
# pander(alpha_table_Herbicide, caption = "Fungi: Herbicide effect in experiment 1")
# 
# 
# ## Experiment 2 ##
# 
# ### pair-wise tests
# alpha_Herbicide_aov_tukey <- emmeans(exp2_falpha_all_aov, "Herbicide")  # library(emmeans)
# alpha_Herbicide_aov_tukey_letters <- cld(alpha_Herbicide_aov_tukey, Letter="abcdefghi", alpha=0.05)
# row.names(alpha_Herbicide_aov_tukey_letters) <- alpha_Herbicide_aov_tukey_letters[,1]
# alpha_Herbicide_aov_tukey_letters <- alpha_Herbicide_aov_tukey_letters[levels(exp2_fDESIGN$Herbicide),]
# # means
# alpha_Herbicide_means <- aggregate(exp2_alpha_diversity_fDAT_summary$asv_diversity,list(exp2_alpha_diversity_fDAT_summary$Herbicide), mean)
# row.names(alpha_Herbicide_means) <- alpha_Herbicide_means[,1]
# alpha_Herbicide_means <- alpha_Herbicide_means[levels(exp2_fDESIGN$Herbicide),]
# # table
# alpha_table_Herbicide <- cbind(Shannon=alpha_Herbicide_means[,2], alpha_Herbicide_aov_tukey_letters[,2:7])
# alpha_table_Herbicide <- alpha_table_Herbicide[,c(1,7)]
# 
# pander(alpha_table_Herbicide, caption = "Fungi: Herbicide effect in experiment 2")

```


```{r plot a-div herbicides, echo=F, message=F, warning=F, eval=T, fig.height=5, fig.width=8}

# ### Figure 4.3 | Shannon diversity - Herbicide
# \vspace{5mm}

# ### BACTERIA
# 
# #data frame for letters
# letters <- data.frame(label=c(rep("a",6)),
#                       Herbicide=rep(c("ctr", "gly", "soil"), 2),
#                       Experiment=rep(c("Experiment 1","Experiment 2"),each=3),
#                       x=c(rep(c(1:3),2)),
#                       y=c(800,800,800,650,650,650))
# 
# alpha_diversity_bDAT_summary$Time <- as.factor(alpha_diversity_bDAT_summary$Time)
# levels(alpha_diversity_bDAT_summary$Herbicide) <- c("ctr", "gly", "tb")
# levels(alpha_diversity_bDAT_summary$Experiment) <- c("Experiment 1", "Experiment 2")
# 
# 
# #plot
# bac_plot <- ggplot(alpha_diversity_bDAT_summary, aes(y=asv_diversity, x=Herbicide, fill=Herbicide)) +
#               geom_boxplot(position=position_dodge2(width=0.75, preserve="single"),outlier.colour = NA) +
#               geom_jitter(size=1, position=position_jitterdodge(jitter.width=0.3,dodge.width = 0.75), alpha=0.3)+
#               theme_bw() +
#               theme(legend.position="none") +
#               ylab("Shanon Diversity")+
#               xlab("")+
#               scale_fill_manual(values=level_cols_Herbicide) +
#               facet_grid(.~ Experiment) +
#               labs(title = "Bacteria: Alpha diversity", subtitle = "alpha diversity ~ Herbicide * Application * Time + Plate")+
#               theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.subtitle=element_text(size=10))+
#               geom_text(data = letters, aes(x = x, y = y, label = label, Herbicide=Herbicide),size=5, col="dimgrey")
# 
# 
# ### FUNGI
# 
# #data frame for letters
# letters <- data.frame(label=c(rep("a",6)),
#                       Herbicide=rep(c("ctr", "gly", "soil"), 2),
#                       Experiment=rep(c("Experiment 1","Experiment 2"),each=3),
#                       x=c(rep(c(1:3),2)),
#                       y=c(58,58,58,58,58,58))
# 
# alpha_diversity_fDAT_summary$Time <- as.factor(alpha_diversity_fDAT_summary$Time)
# levels(alpha_diversity_fDAT_summary$Herbicide) <- c("ctr", "gly", "tb")
# levels(alpha_diversity_fDAT_summary$Experiment) <- c("Experiment 1", "Experiment 2")
# 
# 
# #plot
# fun_plot <- ggplot(alpha_diversity_fDAT_summary, aes(y=asv_diversity, x=Herbicide, fill=Herbicide)) +
#               geom_boxplot(position=position_dodge2(width=0.75, preserve="single"),outlier.colour = NA) +
#               geom_jitter(size=1, position=position_jitterdodge(jitter.width=0.3,dodge.width = 0.75), alpha=0.3)+
#               theme_bw() +
#               theme(legend.position="none") +
#               ylab("Shanon Diversity")+
#               xlab("")+
#               scale_fill_manual(values=level_cols_Herbicide) +
#               facet_grid(.~ Experiment) +
#               labs(title = "Bacteria: Alpha diversity", subtitle = "alpha diversity ~ Herbicide * Application * Time + Plate")+
#               theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.subtitle=element_text(size=10))+
#               geom_text(data = letters, aes(x = x, y = y, label = label, Herbicide=Herbicide),size=5, col="dimgrey")
# 
# 
# 
# ### PLOT
# 
# #combine the two plots (library cowplot)
# fig4.1 <- plot_grid(bac_plot + theme(legend.position="none"),
#                fun_plot + theme(legend.position="none") + labs(colour=""),
#                labels=c("A","B"),
#                align = "hv",
#                nrow=1)
# 
# #add legend
# legend <- get_legend(bac_plot + theme(legend.position="bottom") + labs(fill="Application"))
# fig4.1 <- plot_grid(fig4.1, legend, nrow=2, rel_heights=c(1, .1))
# 
# #print
# #png("figures/fig4.1.png",6400, 4000, res = 600)
# fig4.1
# #dev.off()

# \vspace{5mm}
# **Conclusion:** As expected, the alpha diversity was not affected by the different herbicides.

```

\vspace{5mm}

## Conclusion alpha diversity 

A inconsistent effect on the fungal alpha diversity was found between the different herbicide applications. The alpha diversity of bacteria and fungi did not changed between water control, glyphosate and terbuthylazine.

```{r export RDA files, , echo=F, message=F, warning=F}

# create directory
dir.create("interim")
dir.create("interim/03_Alpha_Diversity")

## set output directory
setwd("interim/03_Alpha_Diversity")

#save objects needed in the following scripts as RDA
saveRDS(alpha_diversity_bDAT_summary, "alpha_diversity_bDAT_summary.RDS")
saveRDS(alpha_diversity_fDAT_summary, "alpha_diversity_fDAT_summary.RDS")

```

\pagebreak
