---
title: "Herbicide - 05 BX Taxa Response"
author: "Jan Waelchli"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---
   
```{r setup, include=FALSE}

##clear the object from memory
rm(list=ls())

#knitr settings
knitr::opts_chunk$set(echo=TRUE, fig.align="center")
options(tinytex.verbose = TRUE)

#set seed
set.seed(100)

#paths
paths <- list(wd = c("/Users/j.waechli/Desktop/Jan_Waelchli/01_Projects/01_Support/Vero/02_ASV_herbicide_trt/5_stats/06_stats_for_paper4"),
              functions="../functions/"
          )

## set source
setwd(paths$wd)

## load functions
source(paste0(paths$functions,"functions.R"))
functions(path="../functions/")

## installs (if necessary) and loads libraries
libraries()
library(ggvenn)

```

```{r paths, echo=F, message=F, warning=F}

# Import

#set wd to RDS files
setwd("interim/01_Import_Normalization")
#Import RDS files
bDESIGN <- readRDS("bDESIGN.RDS")
bDAT <- readRDS("bDAT.RDS")
bDAT_rare <- readRDS("bDAT_rare.RDS")
bTAX <- readRDS("bTAX.RDS")
b_rare <- readRDS("b_rare.RDS")

fDESIGN <- readRDS("fDESIGN.RDS")
fDAT <- readRDS("fDAT.RDS")
fDAT_rare <- readRDS("fDAT_rare.RDS")
fTAX <- readRDS("fTAX.RDS")
f_rare <- readRDS("f_rare.RDS")

level_cols_Herbicide <- readRDS("level_cols_Herbicide.RDS")
level_cols_Spraying <- readRDS("level_cols_Spraying.RDS")

bPHYSEQ <- readRDS("../02_Taxa_Analysis/bPHYSEQ.RDS")
fPHYSEQ <- readRDS("../02_Taxa_Analysis/fPHYSEQ.RDS")

```

\pagebreak

# Taxa Response

Did a core of microbial taxa respond to herbicides? We searched sensitive ASVs – ASVs being differential abundant between two treatments – using edgeR (Robinson et al., 2010).  We followed the same structure as before and answered the 2 questions:

* **Q1: Are there sensitive ASVs between the spraying applications?**
* **Q2: Are there sensitive ASVs between the herbicide treatments?**

We answered the question by using four different tools to measure differential abundances - aldex2, acombc, maaslin2 and metagenomeSeq - and predict ASVs to be different if they were detected by 3 or more tools.

```{r prepro, echo=F, message=F, warning=F}

preprocess <- function(DESIGN, DAT, sample_var){
  
  #preprocess (from ancom.R function file)
  #out_cut: detect outlayers
  #zero_cut: removes ASVs must be present in 90% of samples
  prepro <- feature_table_pre_process(feature_table=DAT, meta_data=DESIGN, 
                                       sample_var=sample_var, group_var=NULL, 
                                       out_cut=0.5, zero_cut=0.9,
                                       lib_cut=0, neg_lb=F)
  #overwrite
  feature_table <-  prepro$feature_table # Preprocessed feature table
  meta_data <-  prepro$meta_data # Preprocessed metadata
  return(list(feature_table, meta_data))
}

```

```{r function herbicide timepoint, warning=F, echo=F}

get_Herbicide_Timepoint_Experiment <- function(DESIGN, DAT, TAX, Herbicide, Timepoint, Experiment=c(1,2)){
  
  #DESIGN
  #subsetting to compartment samples
  DESIGN_Herbicide_Timepoint_Experiment <- DESIGN [DESIGN$Timepoint %in% Timepoint & DESIGN$Herbicide %in% Herbicide & DESIGN$Experiment %in% Experiment,]
  DESIGN_Herbicide_Timepoint_Experiment <- droplevels(DESIGN_Herbicide_Timepoint_Experiment)
  
  #ASV table
  DAT_Herbicide_Timepoint_Experiment <- DAT[,DESIGN_Herbicide_Timepoint_Experiment$Label]
  
  #taxa table
  TAX_Herbicide_Timepoint_Experiment <- TAX[rownames(TAX) %in% rownames(DAT_Herbicide_Timepoint_Experiment),]
  
  return(list(DESIGN_Herbicide_Timepoint_Experiment, DAT_Herbicide_Timepoint_Experiment, TAX_Herbicide_Timepoint_Experiment))
}

```

```{r function sens Spraying, warning=F, echo=F}

sensitive_to_spraying <- function(DESIGN, DAT, TAX, Herbicide, Timepoint, Experiment){
  
  #subset datasets
  temp <-   get_Herbicide_Timepoint_Experiment(DESIGN, DAT, TAX, Herbicide, Timepoint, Experiment)
  DES_sub <- temp[[1]]
  DAT_sub <- temp[[2]] %>% as.data.frame()
  TAX_sub <- temp[[3]]
  rm(temp)
  
  #preprocess
  temp <- preprocess(DES_sub, DAT_sub, "Label")
  feature_table <- temp[[1]]; meta_data <- temp[[2]]; rm(temp)
  
  #search differential abundance
  temp <- run_all_DAA(feature_table, meta_data, group = "Spraying", DAT_sample_names =  "Label", reference = "weeds,soil")
  summ <- temp[[1]]
  rm(temp)

  #relative abundance
  DAT_sub_rel <- (DAT_sub/colSums(DAT_sub)) * 100
  DAT_sub_rel <- as.data.frame(DAT_sub_rel[ rowSums(DAT_sub_rel) > 0,])
  
  #means & standard error
  means <- aggregate(t(DAT_sub_rel), list(DES_sub$Spraying), mean) %>%
    column_to_rownames("Group.1") %>%
    t() %>% as.data.frame() %>%
    rownames_to_column("ASV")
  
  se <- aggregate(t(DAT_sub_rel), list(DES_sub$Spraying), se) %>%
    column_to_rownames("Group.1") %>%
    t() %>% as.data.frame() %>%
    rownames_to_column("ASV")
  
  #sort by abundance of weeds samples
  order <- order(-means$`weeds`)
  means <- means[order,]
  se <- se[order,]
  
  summary <- pivot_longer(means, !(ASV), names_to="Spraying", values_to="means")
  summary$se <- pivot_longer(se, !(ASV), names_to="Spraying", values_to="se") %>% pull(se)
  summary$Spraying <- factor(summary$Spraying, levels=c("weeds", "soil"))
  summary$ASV <- factor(summary$ASV, levels=unique(summary$ASV)) #keep the sorted order of ASVs
  
  #get score
  summary$score <- 0
    for(a in unique(summary$ASV)){
      score <- summ %>% filter(ASV == a) %>% pull(score)
      if(length(score)==1){summary$score[(summary$ASV==a)] <- score}
    }
  
  #add if different abundant (score > 2)
  summary$sensitive <- ifelse(summary$score > 2, "sensitive", "not_sensitive")
  summary$sensitive <- as.factor(summary$sensitive)
  
  #add abundace in trt compared to ctrl
  summary$abu_trt <- "unchanged"
  for(a in unique(summary$ASV)){
    mean_ctrl <- summary$means[(summary$ASV==a) & (summary$Spraying=="weeds")]
    mean_trt <- summary$means[(summary$ASV==a) & (summary$Spraying=="soil")]
    abu_sens <- summary$sensitive[(summary$ASV==a) & (summary$Spraying=="soil")]
    if(abu_sens == "sensitive"){
      trt_dir <- ifelse(mean_ctrl > mean_trt, "higher in weeds", "higher in soil")
      summary$abu_trt[(summary$ASV==a)] <- trt_dir
      }
  }
  
  #add phyla and class
  summary$phylum <- NA
  summary$class <- NA
  summary$family <- NA
  for(a in unique(summary$ASV)){
    summary$phylum[summary$ASV == a] <- as.character(TAX_sub$phylum[TAX_sub$ASV_ID == a])
    summary$class[summary$ASV == a] <- as.character(TAX_sub$class[TAX_sub$ASV_ID == a])
    summary$family[summary$ASV == a] <- as.character(TAX_sub$family[TAX_sub$ASV_ID == a])
  }
  
  return(list(summ, summary))
  
}

```

\pagebreak

## Spraying effect

**Q1: Are there sensitive ASVs between the spraying applications?**  
The herbicide was either sprayed directly on soil or on weeds. We checked for different abundant ASVs after spraying herbicide.

### Experiment 1

```{r DAA spraying, warning=F, echo=F}

#BACTERIA
temp <- sensitive_to_spraying(bDESIGN, bDAT_rare, bTAX, c("ctr","gly","tb"), c(2,5), Experiment = 1)
bsumm_exp1 <- temp[[1]]; bsummary_exp1 <- temp[[2]]; rm(temp)

temp <- sensitive_to_spraying(bDESIGN, bDAT_rare, bTAX, c("ctr","gly","tb"), c(2,5), Experiment = 2)
bsumm_exp2 <- temp[[1]]; bsummary_exp2 <- temp[[2]]; rm(temp)

#FUNGI
temp <- sensitive_to_spraying(fDESIGN, fDAT_rare, fTAX, c("ctr","gly","tb"), c(2,5), Experiment = 1)
fsumm_exp1 <- temp[[1]]; fsummary_exp1 <- temp[[2]]; rm(temp)

temp <- sensitive_to_spraying(fDESIGN, fDAT_rare, fTAX, c("ctr","gly","tb"), c(2,5), Experiment = 2)
fsumm_exp2 <- temp[[1]]; fsummary_exp2 <- temp[[2]]; rm(temp)

```

```{r DAA spraying output, warning=F, echo=F}

#BACTERIA

#exp1

#nbm sens
bchanged_abu_exp1 <- sum(bsummary_exp1$means[bsummary_exp1$sensitive=="sensitive" & bsummary_exp1$Spraying=="weeds"])
bchanged_abu_exp1 <- paste0(round(bchanged_abu_exp1,2), "%")
bsens_exp1 <- table(bsummary_exp1$abu_trt[bsummary_exp1$ASV %in% bsumm_exp1$ASV & bsummary_exp1$Spraying=="weeds"])
bsens_exp1 <- c(bsens_exp1, "0"); names(bsens_exp1)[2] <- "higher in weeds"
bsens_exp1 <- c(bsens_exp1, "0"); names(bsens_exp1)[3] <- "higher in soil"
bsens_exp1 <- bsens_exp1[c(2,1,3)]
#pander(bsens_exp1, caption="Bacteria exp1: Spraying effect")

#sensitive ASVs
bsensitive_exp1 <- pivot_wider(bsummary_exp1,id_cols = !(c(se, score)), names_from = "Spraying", values_from = c("means", "abu_trt", "sensitive"))
bsensitive_exp1 <- bsensitive_exp1[bsensitive_exp1$sensitive_soil =="sensitive",] #sensitive
bsensitive_exp1 <- bsensitive_exp1[,colnames(bsensitive_exp1) %in% c("ASV","means_weeds","means_soil", "phylum", "class")] #column to pander
bsensitive_exp1$means_weeds <- paste0(round(bsensitive_exp1$means_weeds,3),"%")
bsensitive_exp1$means_soil <- paste0(round(bsensitive_exp1$means_soil,3),"%")
colnames(bsensitive_exp1)[4:5] <- c("abu weeds", "abu soil")
#pander(bsensitive_exp1, caption="Bacteria: sensitive ASVs in exp1")

#exp2

#nbm sens
bchanged_abu_exp2 <- sum(bsummary_exp2$means[bsummary_exp2$sensitive=="sensitive" & bsummary_exp2$Spraying=="weeds"])
bchanged_abu_exp2 <- paste0(round(bchanged_abu_exp2,2), "%")
bsens_exp2 <- table(bsummary_exp2$abu_trt[bsummary_exp2$ASV %in% bsumm_exp2$ASV & bsummary_exp2$Spraying=="weeds"])
bsens_exp2 <- c(bsens_exp2, "0"); names(bsens_exp2)[2] <- "higher in weeds"
bsens_exp2 <- c(bsens_exp2, "0"); names(bsens_exp2)[3] <- "higher in soil"
bsens_exp2 <- bsens_exp2[c(2,1,3)]
#pander(bsens_exp2, caption="Bacteria exp2: Spraying effect")

#sensitive ASVs
bsensitive_exp2 <- pivot_wider(bsummary_exp2,id_cols = !(c(se, score)), names_from = "Spraying", values_from = c("means", "abu_trt", "sensitive"))
bsensitive_exp2 <- bsensitive_exp2[bsensitive_exp2$sensitive_soil =="sensitive",] #sensitive
bsensitive_exp2 <- bsensitive_exp2[,colnames(bsensitive_exp2) %in% c("ASV","means_weeds","means_soil", "phylum", "class")] #column to pander
bsensitive_exp2$means_weeds <- paste0(round(bsensitive_exp2$means_weeds,3),"%")
bsensitive_exp2$means_soil <- paste0(round(bsensitive_exp2$means_soil,3),"%")
colnames(bsensitive_exp2)[4:5] <- c("abu weeds", "abu soil")
#pander(bsensitive_exp2, caption="Bacteria: sensitive ASVs in exp2")


#FUNGI

#exp1

#nbm sens
fchanged_abu_exp1 <- sum(fsummary_exp1$means[fsummary_exp1$sensitive=="sensitive" & fsummary_exp1$Spraying=="weeds"])
fchanged_abu_exp1 <- paste0(round(fchanged_abu_exp1,2), "%")
fsens_exp1 <- table(fsummary_exp1$abu_trt[fsummary_exp1$ASV %in% fsumm_exp1$ASV & fsummary_exp1$Spraying=="weeds"])
fsens_exp1 <- c(fsens_exp1, "0"); names(fsens_exp1)[2] <- "higher in weeds"
fsens_exp1 <- c(fsens_exp1, "0"); names(fsens_exp1)[3] <- "higher in soil"
fsens_exp1 <- fsens_exp1[c(2,1,3)]
#pander(fsens_exp1, caption="Fungi exp1: Spraying effect")

#sensitive ASVs
fsensitive_exp1 <- pivot_wider(fsummary_exp1,id_cols = !(c(se, score)), names_from = "Spraying", values_from = c("means", "abu_trt", "sensitive"))
fsensitive_exp1 <- fsensitive_exp1[fsensitive_exp1$sensitive_soil =="sensitive",] #sensitive
fsensitive_exp1 <- fsensitive_exp1[,colnames(fsensitive_exp1) %in% c("ASV","means_weeds","means_soil", "phylum", "class")] #column to pander
fsensitive_exp1$means_weeds <- paste0(round(fsensitive_exp1$means_weeds,3),"%")
fsensitive_exp1$means_soil <- paste0(round(fsensitive_exp1$means_soil,3),"%")
colnames(fsensitive_exp1)[4:5] <- c("abu weeds", "abu soil")
#pander(fsensitive_exp1, caption="Fungi: sensitive ASVs in exp1")

#exp2

#nbm sens
fchanged_abu_exp2 <- sum(fsummary_exp2$means[fsummary_exp2$sensitive=="sensitive" & fsummary_exp2$Spraying=="weeds"])
fchanged_abu_exp2 <- paste0(round(fchanged_abu_exp2,2), "%")
fsens_exp2 <- table(fsummary_exp2$abu_trt[fsummary_exp2$ASV %in% fsumm_exp2$ASV & fsummary_exp2$Spraying=="weeds"])
fsens_exp2 <- c(fsens_exp2, "0"); names(fsens_exp2)[2] <- "higher in weeds"
fsens_exp2 <- c(fsens_exp2, "0"); names(fsens_exp2)[3] <- "higher in soil"
fsens_exp2 <- fsens_exp2[c(2,1,3)]
#pander(fsens_exp2, caption="Fungi exp2: Spraying effect")

#sensitive ASVs
fsensitive_exp2 <- pivot_wider(fsummary_exp2,id_cols = !(c(se, score)), names_from = "Spraying", values_from = c("means", "abu_trt", "sensitive"))
fsensitive_exp2 <- fsensitive_exp2[fsensitive_exp2$sensitive_soil =="sensitive",] #sensitive
fsensitive_exp2 <- fsensitive_exp2[,colnames(fsensitive_exp2) %in% c("ASV","means_weeds","means_soil", "phylum", "class")] #column to pander
fsensitive_exp2$means_weeds <- paste0(round(fsensitive_exp2$means_weeds,3),"%")
fsensitive_exp2$means_soil <- paste0(round(fsensitive_exp2$means_soil,3),"%")
colnames(fsensitive_exp2)[4:5] <- c("abu weeds", "abu soil")
#pander(fsensitive_exp2, caption="Fungi: sensitive ASVs in exp2")



#combine into one df
bsens <- rbind(bsens_exp1, bsens_exp2, fsens_exp1, fsens_exp2)
bsens <- cbind(taxa=c("bacteria","bacteria", "fungi", "fungi"), experiment=paste("Experiment",c(1:2,1:2)), bsens)
rownames(bsens) <- NULL #remove rownames
#add total abu
abu_weeds <- c(gsub("%", "", bsensitive_exp1$`abu weeds`) %>% as.numeric() %>% sum() %>% paste0("%"),
              gsub("%", "", bsensitive_exp2$`abu weeds`) %>% as.numeric() %>% sum() %>% paste0("%"),
              gsub("%", "", fsensitive_exp1$`abu weeds`) %>% as.numeric() %>% sum() %>% paste0("%"),
              gsub("%", "", fsensitive_exp2$`abu weeds`) %>% as.numeric() %>% sum() %>% paste0("%"))
bsens <- cbind(bsens, `rel. abu. of sens. ASVs`=abu_weeds)
pander(bsens, caption = "Spraying effect")


```

### Figure 6.1 | Bacteria: MA & Venn - Spraying

\vspace{5mm}


#######################
        I'M HERE
#######################




```{r MA and venn bacteria spraying, echo=F, warning=F, message=F, eval=T, fig.height=8, fig.width=9}

#colors
col_weeds_up <- level_cols_Spraying[names(level_cols_Spraying) == "weeds"] %>% as.character()
col_soil_up <- level_cols_Spraying[names(level_cols_Spraying) == "soil"] %>% as.character()

#create bacteria df
df1 <- bEDGER_fit_lrt_topTags_exp1[["table"]]
df1$experiment <- "Experiment 1"
df2 <- bEDGER_fit_lrt_topTags_exp2[["table"]]
df2$experiment <- "Experiment 2"
df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "lower in ctr"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]

#plot
fig6.1.1 <- ggplot(df, aes(x=logCPM, y=logFC))+
              geom_point(aes(col=cols), alpha=0.8, size=2)+
              scale_color_manual(values=c(col_weeds_up,col_soil_up, "dimgrey"))+
              facet_wrap(~ experiment)+
              theme_bw()+
              ylab("log2 fold change")+
              xlab("log2 mean abundance")


#VENN

#higher in ctr
#sensitive bacteria
exp1_ASV_higher <- unique(bEDGER_fit_lrt_table_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]>0])
exp2_ASV_higher <- unique(bEDGER_fit_lrt_table_exp2[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]>0])
list_venn <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
bCRA_higher_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly","tb"), c(2,5), 2, bchanged_exp1[bchanged_exp1$logFC > 0,])[[1]]
bCRA_higher_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly","tb"), c(2,5), 2, bchanged_exp2[bchanged_exp2$logFC > 0,])[[1]]
#labels
labels <- c(paste0("exp 1\ncRA: ", round(bCRA_higher_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_higher_exp2, 2), "%"))
names(list_venn) <- labels
#plot
bvenn_h <- ggvenn(list_venn, fill_color = c(col_weeds_up, col_weeds_up), stroke_size = 0.5, set_name_size = 4, show_percentage = F) +
  facet_wrap(~"higher in weeds")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines
  

list_venn1 <- list_venn

#lower in ctr
#sensitive bacteria
exp1_ASV_lower <- unique(bEDGER_fit_lrt_table_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(bEDGER_fit_lrt_table_exp2[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]<0])
list_venn <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
bCRA_lower_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly","tb"), c(2,5), 2, bchanged_exp1[bchanged_exp1$logFC < 0,])[[1]]
bCRA_lower_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly","tb"), c(2,5), 2, bchanged_exp2[bchanged_exp2$logFC < 0,])[[1]]
#labels
labels <- c(paste0("exp 1\ncRA: ", round(bCRA_lower_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_lower_exp2, 2), "%"))
names(list_venn) <- labels
#plot
bvenn_l <- ggvenn(list_venn, fill_color = c(col_soil_up, col_soil_up), stroke_size = 0.5, set_name_size = 4, show_percentage = F) +
  facet_wrap(~"higher in soil")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines

list_venn2 <- list_venn

#combine plots
fig6.1.2 <- plot_grid(bvenn_h + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 bvenn_l+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=1)

fig6.1 <- plot_grid(fig6.1.1 + theme(legend.position="none"),
                 fig6.1.2+ theme(legend.position="none") + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=2)

fig6.1

#saveRDS(fig6.1, "paper_figs/RDS/fig6.1.RDS")

#list_venn <- c(list_venn1, list_venn2)
# bvenn <- ggvenn(list_venn, fill_color = c(col_weeds_up, col_weeds_up, col_soil_up, col_soil_up), stroke_size = 0.5, set_name_size = 4, show_percentage = F) +
#   facet_wrap(~"Different abundant ASVs")+ #to set title
#   theme_bw()+ #to get grey title box
#   theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
#         axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
#         line = element_blank()) #to remove axes and axes lines

```

\vspace{5mm}

### Figure 6.2 | Fungi: MA & Venn - Spraying

\vspace{5mm}

```{r MA and venn fungi spraying, echo=F, warning=F, message=F, eval=T, fig.height=8, fig.width=9}

#create fungi df
df1 <- fEDGER_fit_lrt_topTags_exp1[["table"]]
df1$experiment <- "Experiment 1"
df2 <- fEDGER_fit_lrt_topTags_exp2[["table"]]
df2$experiment <- "Experiment 2"
df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "lower in ctr"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]


fig6.2.1 <- ggplot(df, aes(x=logCPM, y=logFC))+
              geom_point(aes(col=cols), alpha=0.8, size=2)+
              scale_color_manual(values=c(col_weeds_up,col_soil_up, "dimgrey"))+
              facet_wrap(~ experiment)+
              theme_bw()+
              ylab("log2 fold change")+
              xlab("log2 mean abundance")


#VENN

#higher in ctr
#sensitive fungi
exp1_ASV_higher <- unique(fEDGER_fit_lrt_table_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]>0])
exp2_ASV_lower <- unique(fEDGER_fit_lrt_table_exp2[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]>0])
list_venn <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
fCRA_higher_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly","tb"), c(2,5), 2, fchanged_exp1[fchanged_exp1$logFC > 0,])[[1]]
fCRA_higher_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly","tb"), c(2,5), 2, fchanged_exp2[fchanged_exp2$logFC > 0,])[[1]]
#labels
labels <- c(paste0("exp 1\ncRA: ", round(fCRA_higher_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_higher_exp2, 2), "%"))
names(list_venn) <- labels
#plot
fvenn_h <- ggvenn(list_venn, fill_color = c(col_weeds_up, col_weeds_up), stroke_size = 0.5, set_name_size = 4, show_percentage = F) +
  facet_wrap(~"higher in weeds")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines
  

#lower in ctr
#sensitive fungi
exp1_ASV_lower <- unique(fEDGER_fit_lrt_table_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(fEDGER_fit_lrt_table_exp2[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_exp1[["table"]][["logFC"]]<0])
list_venn <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
fCRA_lower_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly","tb"), c(2,5), 2, fchanged_exp1[fchanged_exp1$logFC < 0,])[[1]]
fCRA_lower_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly","tb"), c(2,5), 2, fchanged_exp2[fchanged_exp2$logFC < 0,])[[1]]
#labels
labels <- c(paste0("exp 1\ncRA: ", round(fCRA_lower_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_lower_exp2, 2), "%"))
names(list_venn) <- labels
#plot
fvenn_l <- ggvenn(list_venn, fill_color = c(col_soil_up, col_soil_up), stroke_size = 0.5, set_name_size = 4, show_percentage = F) +
  facet_wrap(~"higher in soil")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines

#combine plots
fig6.2.2 <- plot_grid(fvenn_h + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 fvenn_l+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=1)

fig6.2 <- plot_grid(fig6.2.1 + theme(legend.position="none"),
                 fig6.2.2+ theme(legend.position="none") + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=2)

fig6.2

#saveRDS(fig6.2, "paper_figs/RDS/fig6.2.RDS")

```


**Conclusion:** Different spraying targets altered bacterial and fungal community members in their relative abundance. The majority of sensitive ASVs are not abundant in the rhizosphere, that's why only small percentages were changed due to spraying herbicides. Most of the changed ASVs were only sensitive either in experiment 1 or experiment 2.


## Herbicide effect

**Q2: Are there sensitive ASVs between the herbicide treatments?**  
We treated samples either with water (control), glyphosate or terbuthylazine. We checked for both experiments for differences in the ASV abundance by comparing the herbicides against the water control.

```{r function Herbicide, warning=F, echo=F}

sensitive_to_herbicide <- function(DESIGN, DAT, TAX, Herbicide, Timepoint, Experiment, threshold.filter=7, species="bacteria"){
  
  #subset datasets
  temp <-   get_Herbicide_Timepoint_Experiment(DESIGN, DAT, TAX, Herbicide, Timepoint, Experiment)
  DES_sub <- temp[[1]]
  DAT_sub <- temp[[2]]
  TAX_sub <- temp[[3]]
  rm(temp)
  
  ### edgeR statistics
  ## Preparing EdgeR object, filtering and normalization (TMM) using
  # custom function edgeR.object.TMM() - source("functions/fun_norm_edgeR_obj.R") 
  EDGER <- edgeR.object.TMM(dat=DAT_sub, threshold.filter=threshold.filter, groups=DES_sub$Herbicide, tax=TAX_sub)
  ## fit model, estimate dispersion (on edgeR object) using
  # custom function fit.model.edgeR() - source("functions/fun_norm_edgeR_obj.R") 
  EDGER_fit <- fit.model.edgeR(EDGER, DES_sub$Herbicide)

  #glmLRT - weeds vs soil
  if("gly" %in% Herbicide){contrasts <- makeContrasts(gly - ctr, levels=DES_sub$Herbicide)}
  if("tb" %in% Herbicide){contrasts <- makeContrasts(tb - ctr, levels=DES_sub$Herbicide)}
  EDGER_fit_lrt <- glmLRT(EDGER_fit, contrast=contrasts)
  EDGER_fit_lrt_topTags <- topTags(EDGER_fit_lrt, n=Inf, p.value=1)  # all stat tests
  EDGER_fit_lrt_table <- topTags(EDGER_fit_lrt, n=nrow(EDGER_fit$counts), adjust.method="BH", sort.by="PValue", p.value=0.05)
  EDGER_fit_lrt_table_ASVs <- rownames(EDGER_fit_lrt_table)
  EDGER_fit_lrt_table_ASVs_percent <- round(100 / nrow(EDGER_fit$counts) * length(EDGER_fit_lrt_table_ASVs), 1)

  # summary of differentially abundant ASVs
  EDGER_summary <- t(summary(decideTestsDGE(EDGER_fit_lrt, p.value=0.05)))
  if("gly" %in% Herbicide){colnames(EDGER_summary) <- c("lower in gly", "unchanged", "higher in gly")}
  if("tb" %in% Herbicide){colnames(EDGER_summary) <- c("lower in tb", "unchanged", "higher in tb")}
  rownames(EDGER_summary) <- NULL

  # # Identification of differentially abundant ASVs
  lrt_ASVs <- rownames(topTags(EDGER_fit_lrt, n=nrow(DAT_sub), adjust.method="BH", sort.by="PValue", p.value=0.05))
  changed <- cbind(TAX[lrt_ASVs, ], EDGER_fit_lrt_table[["table"]][["logFC"]])
  if(species=="bacteria"){try(colnames(changed)[10] <- "logFC", silent = T)}
  if(species=="fungi"){try(colnames(changed)[11] <- "logFC", silent = T)}

  return(list(EDGER_summary, changed, EDGER_fit_lrt_table_ASVs, EDGER_fit_lrt_topTags, lrt_ASVs, EDGER_fit_lrt_table))

  
}

```

```{r function cRA Herbicide, warning=F, echo=F}

cRA_Herbicide <- function(DESIGN, DAT_rare, Herbicide, Timepoint, Experiment, changed){
  
  ##samples
  samples_hc <- DESIGN$Label[which(DESIGN$Herbicide %in% Herbicide & DESIGN$Timepoint %in% Timepoint & DESIGN$Experiment %in% Experiment)]
  samples_ctr <- DESIGN$Label[which(DESIGN$Herbicide %in% "ctr" & DESIGN$Timepoint %in% Timepoint & DESIGN$Experiment %in% Experiment)]
  
  #subsetting and normalization
  DAT_rare_hc <- DAT_rare[, samples_hc ]
  DAT_rare_hc <- t(t(DAT_rare_hc)/colSums(DAT_rare_hc)) * 100 #normalize (each sample = 100%)
  DAT_rare_hc <- DAT_rare_hc[ rowSums(DAT_rare_hc) > 0,] #remove empty ASV
  
  DAT_rare_ctr <- DAT_rare[, samples_ctr ]
  DAT_rare_ctr <- t(t(DAT_rare_ctr)/colSums(DAT_rare_ctr)) * 100
  DAT_rare_ctr <- DAT_rare_ctr[ rowSums(DAT_rare_ctr) > 0,]
  
  ## calculation of means and SE
  # hc and sorting for rank abundance 
  DAT_rare_hc_MEAN <- apply(DAT_rare_hc, 1, mean) #mean over all samples for each ASVs
  DAT_rare_hc_MEAN <- sort(DAT_rare_hc_MEAN, decr=T)
  
  # ctr
  DAT_rare_ctr_MEAN <- apply(DAT_rare_ctr, 1, mean)[names(DAT_rare_hc_MEAN)]
  
  # summary
  DAT_rare_MEANs <- as.data.frame(cbind(DAT_rare_hc_MEAN, DAT_rare_ctr_MEAN))
  colnames(DAT_rare_MEANs) <- c("hc","ctr")
  
  #changed
  DAT_rare_MEANs$changed <- F
  DAT_rare_MEANs$changed[rownames(DAT_rare_MEANs) %in% rownames(changed)] <- T
  
  #cumulative relative abundance of changed ASVs between hc and ctr
  cRA_hc <- sum(DAT_rare_MEANs$hc[DAT_rare_MEANs$changed])
  cRA_ctr <- sum(DAT_rare_MEANs$ctr[DAT_rare_MEANs$changed])
  
  #sort by ASV rank
  changed$rank <- as.numeric(gsub("ASV", "", changed$ASV_ID))
  changed <- changed[order(changed$rank),]

  DAT_rare_MEANs$rank <- as.numeric(gsub("ASV", "", rownames(DAT_rare_MEANs)))
  DAT_rare_MEANs <- DAT_rare_MEANs[order(DAT_rare_MEANs$rank),] #sort by ASV rank

  #add abundance to changed-data
  changed <- changed[changed$ASV_ID %in% rownames(DAT_rare_MEANs),]#ommit errors if not all ASVs in changed are in DAT_rare
  changed$rAbu <- DAT_rare_MEANs[rownames(DAT_rare_MEANs) %in% changed$ASV_ID, 1]#add abu
  changed <- changed[order(changed$rAbu, decreasing = T),] #sort by abu
  try(changed$rAbu <- paste0(round(changed$rAbu,2), "%"), silent = T)
  
  return(list(cRA_hc, cRA_ctr, changed))

}

```

### Experiment 1

```{r bacteria herbicide exp1, warning=F, echo=F}

#BAC

#gly
temp <-   sensitive_to_herbicide(bDESIGN, bDAT, bTAX, c("ctr","gly"), c(2,5), 1, species = "bacteria")
bEDGER_summary_gly_exp1 <- temp[[1]]
bchanged_gly_exp1 <- temp[[2]]
bEDGER_fit_lrt_table_ASVs_gly_exp1 <- temp[[3]]
bEDGER_fit_lrt_topTags_gly_exp1 <- temp[[4]]
lrt_bASVs_gly_exp1 <- temp[[5]]
bEDGER_fit_lrt_table_gly_exp1 <- temp[[6]]

temp <- cRA_Herbicide(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 1, bchanged_gly_exp1)
bCRA_gly_exp1 <- temp[[1]]
bchanged_gly_exp1 <- temp[[3]]

pander(bEDGER_summary_gly_exp1, justify='left', caption = "Bacteria: Glyphosate effect in experiment 1") 
print(paste0(round(bCRA_gly_exp1, 2), "%", " of the bacterial abundance changed after spraying glyphosate"))
pander(bchanged_gly_exp1[1:min(10, nrow(bchanged_gly_exp1)), c(2,3,10,12)], caption="The top abundant ASVs")

#tb
temp <-   sensitive_to_herbicide(bDESIGN, bDAT, bTAX, c("ctr","tb"), c(2,5), 1, species = "bacteria")
bEDGER_summary_tb_exp1 <- temp[[1]]
bchanged_tb_exp1 <- temp[[2]]
bEDGER_fit_lrt_table_ASVs_tb_exp1 <- temp[[3]]
bEDGER_fit_lrt_topTags_tb_exp1 <- temp[[4]]
lrt_bASVs_tb_exp1 <- temp[[5]]
bEDGER_fit_lrt_table_tb_exp1 <- temp[[6]]

temp <- cRA_Herbicide(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 1, bchanged_tb_exp1)
bCRA_tb_exp1 <- temp[[1]]
bchanged_tb_exp1 <- temp[[3]]

pander(bEDGER_summary_tb_exp1, justify='left', caption = "Bacteria: Glyphosate effect in experiment 1") 
print(paste0(round(bCRA_tb_exp1, 2), "%", " of the bacterial abundance changed after spraying terbuthylazine"))
pander(bchanged_tb_exp1[1:min(10, nrow(bchanged_tb_exp1)), c(2,3,10,12)], caption="The top abundant ASVs")


#FUN

#gly
temp <-   sensitive_to_herbicide(fDESIGN, fDAT, fTAX, c("ctr","gly"), c(2,5), 1, species = "fungi")
fEDGER_summary_gly_exp1 <- temp[[1]]
fchanged_gly_exp1 <- temp[[2]]
fEDGER_fit_lrt_table_ASVs_gly_exp1 <- temp[[3]]
fEDGER_fit_lrt_topTags_gly_exp1 <- temp[[4]]
lrt_fASVs_gly_exp1 <- temp[[5]]
fEDGER_fit_lrt_table_gly_exp1 <- temp[[6]]

temp <- cRA_Herbicide(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 1, fchanged_gly_exp1)
fCRA_gly_exp1 <- temp[[1]]
fchanged_gly_exp1 <- temp[[3]]

pander(fEDGER_summary_gly_exp1, justify='left', caption = "Fungi: Glyphosate effect in experiment 1") 
print(paste0(round(fCRA_gly_exp1, 2), "%", " of the fungal abundance changed after spraying glyphosate"))
pander(fchanged_gly_exp1[1:min(10, nrow(fchanged_gly_exp1)), c(2,3,10,12)], caption="The top abundant ASVs")

#tb
temp <-   sensitive_to_herbicide(fDESIGN, fDAT, fTAX, c("ctr","tb"), c(2,5), 1, species = "fungi")
fEDGER_summary_tb_exp1 <- temp[[1]]
fchanged_tb_exp1 <- temp[[2]]
fEDGER_fit_lrt_table_ASVs_tb_exp1 <- temp[[3]]
fEDGER_fit_lrt_topTags_tb_exp1 <- temp[[4]]
lrt_fASVs_tb_exp1 <- temp[[5]]
fEDGER_fit_lrt_table_tb_exp1 <- temp[[6]]

temp <- cRA_Herbicide(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 1, fchanged_tb_exp1)
fCRA_tb_exp1 <- temp[[1]]
fchanged_tb_exp1 <- temp[[3]]

pander(fEDGER_summary_tb_exp1, justify='left', caption = "Fungi: Glyphosate effect in experiment 1") 
print(paste0(round(fCRA_tb_exp1, 2), "%", " of the fungal abundance changed after spraying terbuthylazine"))
pander(fchanged_tb_exp1[1:min(10, nrow(fchanged_tb_exp1)), c(2,3,10,12)], caption="The top abundant ASVs")

```

In experiment 1: Glyhphosate changed the abundance of `r paste(length(bEDGER_fit_lrt_table_ASVs_gly_exp1))` bacterial ASVs and `r paste(length(fEDGER_fit_lrt_table_ASVs_gly_exp1))` fungal ASVs. Spraying terbuthylazine led to changes in abundance for `r paste(length(bEDGER_fit_lrt_table_ASVs_tb_exp1))` bacterial ASVs and `r paste(length(fEDGER_fit_lrt_table_ASVs_tb_exp1))` fungal ASVs.

### Experiment 2

```{r bacteria herbicide exp2, warning=F, echo=F}

#BAC

#gly
temp <-   sensitive_to_herbicide(bDESIGN, bDAT, bTAX, c("ctr","gly"), c(2,5), 2, species = "bacteria")
bEDGER_summary_gly_exp2 <- temp[[1]]
fchanged_gly_exp2 <- temp[[2]]
bEDGER_fit_lrt_table_ASVs_gly_exp2 <- temp[[3]]
bEDGER_fit_lrt_topTags_gly_exp2 <- temp[[4]]
lrt_bASVs_gly_exp2 <- temp[[5]]
bEDGER_fit_lrt_table_gly_exp2 <- temp[[6]]

temp <- cRA_Herbicide(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 2, fchanged_gly_exp2)
bCRA_gly_exp2 <- temp[[1]]
fchanged_gly_exp2 <- temp[[3]]

pander(bEDGER_summary_gly_exp2, justify='left', caption = "Bacteria: Glyphosate effect in experiment 2") 
print(paste0(round(bCRA_gly_exp2, 2), "%", " of the bacterial abundance changed after spraying glyphosate"))
pander(fchanged_gly_exp2[1:min(10, nrow(fchanged_gly_exp2)), c(2,3,10,12)], caption="The top abundant ASVs")

#tb
temp <-   sensitive_to_herbicide(bDESIGN, bDAT, bTAX, c("ctr","tb"), c(2,5), 2, species = "bacteria")
bEDGER_summary_tb_exp2 <- temp[[1]]
fchanged_tb_exp2 <- temp[[2]]
bEDGER_fit_lrt_table_ASVs_tb_exp2 <- temp[[3]]
bEDGER_fit_lrt_topTags_tb_exp2 <- temp[[4]]
lrt_bASVs_tb_exp2 <- temp[[5]]
bEDGER_fit_lrt_table_tb_exp2 <- temp[[6]]

temp <- cRA_Herbicide(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 2, fchanged_tb_exp2)
bCRA_tb_exp2 <- temp[[1]]
fchanged_tb_exp2 <- temp[[3]]

pander(bEDGER_summary_tb_exp2, justify='left', caption = "Bacteria: Glyphosate effect in experiment 2") 
print(paste0(round(bCRA_tb_exp2, 2), "%", " of the bacterial abundance changed after spraying terbuthylazine"))
pander(fchanged_tb_exp2[1:min(10, nrow(fchanged_tb_exp2)), c(2,3,10,12)], caption="The top abundant ASVs")


#FUN
#gly
temp <-   sensitive_to_herbicide(fDESIGN, fDAT, fTAX, c("ctr","gly"), c(2,5), 2, species = "fungi")
fEDGER_summary_gly_exp2 <- temp[[1]]
fchanged_gly_exp2 <- temp[[2]]
fEDGER_fit_lrt_table_ASVs_gly_exp2 <- temp[[3]]
fEDGER_fit_lrt_topTags_gly_exp2 <- temp[[4]]
lrt_fASVs_gly_exp2 <- temp[[5]]
fEDGER_fit_lrt_table_gly_exp2 <- temp[[6]]

temp <- cRA_Herbicide(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 2, fchanged_gly_exp2)
fCRA_gly_exp2 <- temp[[1]]
fchanged_gly_exp2 <- temp[[3]]

pander(fEDGER_summary_gly_exp2, justify='left', caption = "Fungi: Glyphosate effect in experiment 2") 
print(paste0(round(fCRA_gly_exp2, 2), "%", " of the fungal abundance changed after spraying glyphosate"))
pander(fchanged_gly_exp2[1:min(10, nrow(fchanged_gly_exp2)), c(2,3,11,13)], caption="The top abundant ASVs")

#tb
temp <-   sensitive_to_herbicide(fDESIGN, fDAT, fTAX, c("ctr","tb"), c(2,5), 2, species = "fungi")
fEDGER_summary_tb_exp2 <- temp[[1]]
fchanged_tb_exp2 <- temp[[2]]
fEDGER_fit_lrt_table_ASVs_tb_exp2 <- temp[[3]]
fEDGER_fit_lrt_topTags_tb_exp2 <- temp[[4]]
lrt_fASVs_tb_exp2 <- temp[[5]]
fEDGER_fit_lrt_table_tb_exp2 <- temp[[6]]

temp <- cRA_Herbicide(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 2, fchanged_tb_exp2)
fCRA_tb_exp2 <- temp[[1]]
fchanged_tb_exp2 <- temp[[3]]

pander(fEDGER_summary_tb_exp2, justify='left', caption = "Fungi: Glyphosate effect in experiment 2") 
print(paste0(round(fCRA_tb_exp2, 2), "%", " of the fungal abundance changed after spraying terbuthylazine"))
pander(fchanged_tb_exp2[1:min(10, nrow(fchanged_tb_exp2)), c(2,3,11,13)], caption="The top abundant ASVs")

```

In experiment 2: Glyhphosate changed the abundance of `r paste(length(bEDGER_fit_lrt_table_ASVs_gly_exp2))` bacterial ASVs and `r paste(length(fEDGER_fit_lrt_table_ASVs_gly_exp2))` fungal ASVs. Spraying terbuthylazine led to changes in abundance for `r paste(length(bEDGER_fit_lrt_table_ASVs_tb_exp2))` bacterial ASVs and `r paste(length(fEDGER_fit_lrt_table_ASVs_tb_exp2))` fungal ASVs.

### Figure 7.1 | Bacteria: MA & Venn - Herbicide

\vspace{5mm}

```{r MA and venn bacteria herbicide, echo=F, warning=F, message=F, eval=T, fig.height=10, fig.width=6}

#colors
col_ctr_up <- level_cols_Herbicide[names(level_cols_Herbicide) == "ctr"] %>% as.character()
col_gly_up <- level_cols_Herbicide[names(level_cols_Herbicide) == "gly"] %>% as.character()
col_tb_up <- level_cols_Herbicide[names(level_cols_Herbicide) == "tb"] %>% as.character()

#gly
#create bacteria df
df1 <- bEDGER_fit_lrt_topTags_gly_exp1[["table"]]
df1$herbicide <- "gly"
df1$experiment <- "Experiment 1"
df2 <- bEDGER_fit_lrt_topTags_gly_exp2[["table"]]
df2$herbicide <- "gly"
df2$experiment <- "Experiment 2"

df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "higher in gly"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]

#plot
fig7.1.1_gly <- ggplot(df, aes(x=logCPM, y=logFC))+
                  geom_point(aes(col=cols), alpha=0.8, size=2)+
                  scale_color_manual(values=c(col_ctr_up,col_gly_up, "dimgrey"))+
                  facet_wrap(~ experiment)+
                  theme_bw()+
                  ylab("log2 fold change")+
                  xlab("log2 mean abundance")

#tb
#create bacteria df
df1 <- bEDGER_fit_lrt_topTags_tb_exp1[["table"]]
df1$herbicide <- "tb"
df1$experiment <- "Experiment 1"
df2 <- bEDGER_fit_lrt_topTags_tb_exp2[["table"]]
df2$herbicide <- "tb"
df2$experiment <- "Experiment 2"

df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "higher in tb"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]

#plot
fig7.1.1_tb <- ggplot(df, aes(x=logCPM, y=logFC))+
                  geom_point(aes(col=cols), alpha=0.8, size=2)+
                  scale_color_manual(values=c(col_ctr_up,col_tb_up, "dimgrey"))+
                  facet_wrap(~ experiment)+
                  theme_bw()+
                  ylab("log2 fold change")+
                  xlab("log2 mean abundance")

#combine MA plots
fig7.1.1 <- plot_grid(fig7.1.1_gly + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 fig7.1.1_tb+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=2)

#VENN

#higher in ctr
#sensitive bacteria gly
exp1_ASV_higher <- unique(bEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]>0])
exp2_ASV_higher <- unique(bEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]>0])
list_venn_gly <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
bCRA_higher_gly_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 1, bchanged_gly_exp1[bchanged_gly_exp1$logFC > 0,])[[1]]
bCRA_higher_gly_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 2, bchanged_gly_exp1[bchanged_gly_exp1$logFC > 0,])[[1]]
#sensitive bacteria tb
exp1_ASV_higher <- unique(bEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]>0])
exp2_ASV_higher <- unique(bEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]>0])
list_venn_tb <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
bCRA_higher_tb_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 1, bchanged_tb_exp1[bchanged_tb_exp1$logFC > 0,])[[1]]
bCRA_higher_tb_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 2, bchanged_tb_exp1[bchanged_tb_exp1$logFC > 0,])[[1]]

#labels
list_venn <- c(list_venn_gly, list_venn_tb)
labels <- c(paste0("exp 1\ncRA: ", round(bCRA_higher_gly_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_higher_gly_exp2, 2), "%"),
            paste0("exp 1\ncRA: ", round(bCRA_higher_tb_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_higher_tb_exp2, 2), "%"))
names(list_venn) <- labels

#plot
bvenn_h <- ggvenn(list_venn, fill_color = c(col_gly_up, col_gly_up, col_tb_up, col_tb_up), stroke_size = 0.5, set_name_size = 2, show_percentage = F) +
  facet_wrap(~"higher in ctr")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines
  

#lower in ctr
#sensitive bacteria gly
exp1_ASV_lower <- unique(bEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(bEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]<0])
list_venn_gly <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
bCRA_lower_gly_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 1, bchanged_gly_exp1[bchanged_gly_exp1$logFC < 0,])[[1]]
bCRA_lower_gly_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","gly"), c(2,5), 2, bchanged_gly_exp1[bchanged_gly_exp1$logFC < 0,])[[1]]
#sensitive bacteria tb
exp1_ASV_lower <- unique(bEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(bEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][bEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]<0])
list_venn_tb <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
bCRA_lower_tb_exp1 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 1, bchanged_tb_exp1[bchanged_tb_exp1$logFC < 0,])[[1]]
bCRA_lower_tb_exp2 <- cRA_Spraying(bDESIGN, bDAT_rare, c("ctr","tb"), c(2,5), 2, bchanged_tb_exp1[bchanged_tb_exp1$logFC < 0,])[[1]]

#labels
list_venn <- c(list_venn_gly, list_venn_tb)
labels <- c(paste0("exp 1\ncRA: ", round(bCRA_lower_gly_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_lower_gly_exp2, 2), "%"),
            paste0("exp 1\ncRA: ", round(bCRA_lower_tb_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(bCRA_lower_tb_exp2, 2), "%"))
names(list_venn) <- labels

#plot
bvenn_l <- ggvenn(list_venn, fill_color = c(col_gly_up, col_gly_up, col_tb_up, col_tb_up), stroke_size = 0.5, set_name_size = 2, show_percentage = F) +
  facet_wrap(~"higher in hc")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines



#combine plots
fig7.1.2 <- plot_grid(bvenn_h + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 bvenn_l+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=1)

fig7.1 <- plot_grid(fig7.1.1 + theme(legend.position="none"),
                 fig7.1.2+ theme(legend.position="none") + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 rel_heights = c(1,.5),
                 nrow=2)

fig7.1

#saveRDS(fig7.1, "paper_figs/RDS/fig7.1.RDS")


```

\vspace{5mm}

### Figure 7.2 | Fungi: MA & Venn - Herbicide

\vspace{5mm}

```{r MA and venn fungi herbicide, echo=F, warning=F, message=F, eval=T, fig.height=10, fig.width=6}

#gly
#create fungi df
df1 <- fEDGER_fit_lrt_topTags_gly_exp1[["table"]]
df1$herbicide <- "gly"
df1$experiment <- "Experiment 1"
df2 <- fEDGER_fit_lrt_topTags_gly_exp2[["table"]]
df2$herbicide <- "gly"
df2$experiment <- "Experiment 2"

df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "higher in gly"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]

#plot
fig7.2.1_gly <- ggplot(df, aes(x=logCPM, y=logFC))+
                  geom_point(aes(col=cols), alpha=0.8, size=2)+
                  scale_color_manual(values=c(col_ctr_up,col_gly_up, "dimgrey"))+
                  facet_wrap(~ experiment)+
                  theme_bw()+
                  ylab("log2 fold change")+
                  xlab("log2 mean abundance")

#tb
#create fungi df
df1 <- fEDGER_fit_lrt_topTags_tb_exp1[["table"]]
df1$herbicide <- "tb"
df1$experiment <- "Experiment 1"
df2 <- fEDGER_fit_lrt_topTags_tb_exp2[["table"]]
df2$herbicide <- "tb"
df2$experiment <- "Experiment 2"

df <- rbind(df1, df2)
rm(df1, df2)

#get up and down-regulated ASVs
df$cols <- "NS"
df$cols[df$FDR < 0.05 & df$logFC > 0] <- "higher in ctr"
df$cols[df$FDR < 0.05 & df$logFC < 0] <- "higher in tb"

#sort to plot sign at the end
df <- df[order(df$FDR, decreasing = T),]

#plot
fig7.2.1_tb <- ggplot(df, aes(x=logCPM, y=logFC))+
                  geom_point(aes(col=cols), alpha=0.8, size=2)+
                  scale_color_manual(values=c(col_ctr_up,col_tb_up, "dimgrey"))+
                  facet_wrap(~ experiment)+
                  theme_bw()+
                  ylab("log2 fold change")+
                  xlab("log2 mean abundance")

#combine MA plots
fig7.2.1 <- plot_grid(fig7.2.1_gly + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 fig7.2.1_tb+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=2)

#VENN

#higher in ctr
#sensitive fungi gly
exp1_ASV_higher <- unique(fEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]>0])
exp2_ASV_higher <- unique(fEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]>0])
list_venn_gly <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
fCRA_higher_gly_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 1, fchanged_gly_exp1[fchanged_gly_exp1$logFC > 0,])[[1]]
fCRA_higher_gly_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 2, fchanged_gly_exp1[fchanged_gly_exp1$logFC > 0,])[[1]]
#sensitive fungi tb
exp1_ASV_higher <- unique(fEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]>0])
exp2_ASV_higher <- unique(fEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]>0])
list_venn_tb <- list(exp1_higher = exp1_ASV_higher, exp2_higher = exp2_ASV_higher)
#relative abundance
fCRA_higher_tb_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 1, fchanged_tb_exp1[fchanged_tb_exp1$logFC > 0,])[[1]]
fCRA_higher_tb_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 2, fchanged_tb_exp1[fchanged_tb_exp1$logFC > 0,])[[1]]

#labels
list_venn <- c(list_venn_gly, list_venn_tb)
labels <- c(paste0("exp 1\ncRA: ", round(fCRA_higher_gly_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_higher_gly_exp2, 2), "%"),
            paste0("exp 1\ncRA: ", round(fCRA_higher_tb_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_higher_tb_exp2, 2), "%"))
names(list_venn) <- labels

#plot
fvenn_h <- ggvenn(list_venn, fill_color = c(col_gly_up, col_gly_up, col_tb_up, col_tb_up), stroke_size = 0.5, set_name_size = 2, show_percentage = F) +
  facet_wrap(~"higher in ctr")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines
  

#lower in ctr
#sensitive fungi gly
exp1_ASV_lower <- unique(fEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(fEDGER_fit_lrt_table_gly_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_gly_exp1[["table"]][["logFC"]]<0])
list_venn_gly <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
fCRA_lower_gly_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 1, fchanged_gly_exp1[fchanged_gly_exp1$logFC < 0,])[[1]]
fCRA_lower_gly_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","gly"), c(2,5), 2, fchanged_gly_exp1[fchanged_gly_exp1$logFC < 0,])[[1]]
#sensitive fungi tb
exp1_ASV_lower <- unique(fEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]<0])
exp2_ASV_lower <- unique(fEDGER_fit_lrt_table_tb_exp1[["table"]][["ASV_ID"]][fEDGER_fit_lrt_table_tb_exp1[["table"]][["logFC"]]<0])
list_venn_tb <- list(exp1_lower = exp1_ASV_lower, exp2_lower = exp2_ASV_lower)
#relative abundance
fCRA_lower_tb_exp1 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 1, fchanged_tb_exp1[fchanged_tb_exp1$logFC < 0,])[[1]]
fCRA_lower_tb_exp2 <- cRA_Spraying(fDESIGN, fDAT_rare, c("ctr","tb"), c(2,5), 2, fchanged_tb_exp1[fchanged_tb_exp1$logFC < 0,])[[1]]

#labels
list_venn <- c(list_venn_gly, list_venn_tb)
labels <- c(paste0("exp 1\ncRA: ", round(fCRA_lower_gly_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_lower_gly_exp2, 2), "%"),
            paste0("exp 1\ncRA: ", round(fCRA_lower_tb_exp1, 2), "%"), 
            paste0("exp 2\ncRA: ", round(fCRA_lower_tb_exp2, 2), "%"))
names(list_venn) <- labels

#plot
fvenn_l <- ggvenn(list_venn, fill_color = c(col_gly_up, col_gly_up, col_tb_up, col_tb_up), stroke_size = 0.5, set_name_size = 2, show_percentage = F) +
  facet_wrap(~"higher in hc")+ #to set title
  theme_bw()+ #to get grey title box
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),
        line = element_blank()) #to remove axes and axes lines



#combine plots
fig7.2.2 <- plot_grid(fvenn_h + theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")),
                 fvenn_l+ theme(legend.position="none", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), "cm")) + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 nrow=1)

fig7.2 <- plot_grid(fig7.2.1 + theme(legend.position="none"),
                 fig7.2.2+ theme(legend.position="none") + labs(colour=""),
                 #labels=c("A","B"),
                 align = "hv",
                 rel_heights = c(1,.5),
                 nrow=2)

fig7.2

#saveRDS(fig7.2, "paper_figs/RDS/fig7.2.RDS")

```


\vspace{5mm}

**Conclusion:** Spraying glyphosate and terbuthylazine led to similar effect. In both experiments were around 30 ASVs sensitive in their abundance to glyphosate and terbuthylazine but non of those ASVs was sensitive in both experiments. We found high abundant bacterial ASVs reacting sensitive to herbicides which was up to 12 % of the bacterial community. Only a few fungal ASVs were changed in their abundance and non of them was sensitive for both herbicides or in both experiments.


**Conclusion taxa response:** *Spraying to different targets did not influence high abundant ASVs. Spraying glyphosate and terbuthylazine caused similar effects. Both altered high abundant bacterial ASVs. Almost no effect was found on fungi. Sensitive ASVs were not consistent between the two experiments*



